\section{Related work}
\label{wasm:rel_works}

There are several successful solutions that leverage Wasm to sandbox
untrusted code~\cite{RLBox, ewasm, sledge}. RLBox~\cite{RLBox} is a
framework that facilitates the isolation of third-party libraries in
pre-existing software. eWASM~\cite{ewasm} optimizes the execution of
Wasm in embedded systems with constrained resources. Sledge~\cite{sledge}
enables efficient Wasm-based serverless execution on the
edge. The use of our approach for restricting access to the file
system within these frameworks can strengthen their security
assurance.

The memory safety guarantees of Wasm depend on the runtime
implementation~\cite{lehmann2020everything}. Hence, Bosamiya et
al.~\cite{bosamiya2022provably} explore the problem of producing
provably safe sandboxes. WaVe~\cite{johnson2022wave} explains that any
interaction with the unsafe interfaces exposed by WASI can
introduce security and safety violations. Thus, the authors proposed
a verified secure runtime system
implementing WASI. However, both works require to redesign the runtime
toolchain, while our solution can be directly integrated into existing
runtimes.
%However, both proposals rely on implementing a new runtime, 
%while our approach integrates with existing runtimes used in the industry.

The academic and industrial communities have investigated the use of eBPF
for the isolation of software~\cite{findlay2020bpfbox, findlay2021bpfcontain,
cilium-repo, falco}.
{\em BPFBox}~\cite{findlay2020bpfbox} and {\em BPFContain}
~\cite{findlay2021bpfcontain} use an eBPF daemon to confine processes and
services.
{\em Cilium}~\cite{cilium-repo} provides eBPF-based networking, observability
and security for container workloads.
{\em Falco}~\cite{falco} enables lightweight threat detection in the
cluster. These solutions highlight the potential of eBPF, and provide
a simple and flexible confinement of system resources. However, they
focus on containers or services, while our solution aims at enforcing
fine-grained per-sandbox policies.

% say something about the match between WASM & eBPF portability ???

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main.tex"
%%% End:
